---
title: "XYZ Retail Analysis"
author: "Anurag Mishra | Eeshita Kumar | Jayadev KP | Jing Xie | Jingwen Pei"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(maps)
library(usmap)
library(grid)

setwd('C:/Users/jaide/OneDrive/Documents/R/HW1/Data')

US <-
  read.csv("us-counties.txt",
           stringsAsFactors = F,
           strip.white = T)
names(US)

US_recent <-
  read.csv("us-counties-recent.txt",
           stringsAsFactors = F,
           strip.white = T)
names(US_recent)


US_total <-
  rbind(
    US %>% select(
      date,
      county,
      state,
      cases,
      cases_avg,
      cases_avg_per_100k,
      deaths,
      deaths_avg,
      deaths_avg_per_100k
    ),
    US_recent %>%  select(
      date,
      county,
      state,
      cases,
      cases_avg,
      cases_avg_per_100k,
      deaths,
      deaths_avg,
      deaths_avg_per_100k
    ) %>% filter(date > "2021-09-29")
  )

rm(US)
rm(US_recent)

states <- map_data("state")

US_total$region <- tolower(US_total$state)

centroid_labels <-
  utils::read.csv(system.file("extdata", paste0("us_", "states", "_centroids.csv"), package = "usmap"),
                  stringsAsFactors = FALSE)


gdp = read.csv("GDP.csv")
gdphigh = read.csv('highlight.csv')


vaccine = read.csv("State Vaccination percentages.csv")



map.df2= merge(states,vaccine, by = "region")
map.df.centers2 <-
  map.df2 %>% group_by(region) %>% summarize(long_c = mean(long, na.rm = T),
                                           lat_c = mean(lat, na.rm = T))
vaccine_lat <-
  merge(
    vaccine,
    map.df.centers2,
    by.x = 'region',
    by.y = 'region',
    all.x = T,
    all.y = F
  )



sales <- read.csv("tsadjustedsales_f.csv")

vec <- c("Q1 2019",
         "Q2 2019",
         "Q3 2019",
         "Q4 2019",
         "Q1 2020",
         "Q2 2020",
         "Q3 2020",
         "Q4 2020",
         "Q1 2021",
         "Q2 2021")



```

## XYZ should identify "opportunity states" with decreasing trends in Covid-19 cases, high vaccination rates, good post-pandemic economic recovery in the retail sector as well as high retail demand in 2021.


### Heatmap of Covid-19 cases across the US from  2020-01-21 to 2021-09-30

Please select a particular date to check case density distribution or press the play button to see the variation in densities across time.

```{r eruptions, echo=FALSE}
inputPanel(
  
  sliderInput("selected_date", label = "Date selected:",
              min = as.Date(min(US_total$date)), max = as.Date(max(US_total$date)), value = as.Date(max(US_total$date)),  timeFormat = "%Y-%m-%d", ticks = F, animate = T)
)

renderPlot({
  US_total_states <-
    US_total %>% filter(date == as.character(input$selected_date)) %>% group_by(state, region) %>% summarize(
      cases_avg_per_100k = sum(cases_avg_per_100k, na.rm = T),
      deaths_avg_per_100k = sum(deaths_avg_per_100k, na.rm = T)
    )
  
  map.df <- merge(states, US_total_states, by = "region", all.x = T)
  map.df <- map.df[order(map.df$order),]
  
  US_total_states_lat <-
    merge(
      US_total_states,
      centroid_labels[, c('full', 'abbr')],
      by.x = 'state',
      by.y = 'full',
      all.x = T,
      all.y = F
    )
  map.df.centers <-
    map.df %>% group_by(state) %>% summarize(long_c = mean(long, na.rm = T),
                                             lat_c = mean(lat, na.rm = T))
  US_total_states_lat <-
    merge(
      US_total_states_lat,
      map.df.centers,
      by.x = 'state',
      by.y = 'state',
      all.x = T,
      all.y = F
    )
  
  
  ggplot(map.df, aes(x = long, y = lat, group = group)) +
    geom_polygon(aes(fill = cases_avg_per_100k)) +
    geom_path() +
    geom_text(
      data = US_total_states_lat,
      aes(
        x = long_c,
        y = lat_c,
        group = NA,
        label = abbr
      ),
      size = 2.5,
      vjust = 0.5,
      hjust = 0.5
    ) +
    scale_fill_gradientn('Cases (per 100k)',
                         colours = rev(heat.colors(10)),
                         na.value = "grey90") +
    coord_map() +
    scale_x_continuous(breaks = NULL) + scale_y_continuous(breaks = NULL) +
    labs(x = "", y = "")  +
    ggtitle(paste0("Heatmap of Covid-19 cases across the country on ", input$selected_date)) +
    theme(panel.background = element_blank(),
          plot.title = element_text(face = 'bold', size = 18))
})
```

### Heatmap of Vaccination rates (percentage of people vaccinated) across the US as of end of Sep 2021
```{r eruptions3, echo=FALSE}

renderPlot({
  ggplot(map.df2, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = Vaccination_perc)) +
  geom_path() +
  geom_text(
    data = vaccine_lat,
    aes(
      x = long_c,
      y = lat_c,
      group = NA,
      label = code
    ),
    size = 2.5,
    vjust = 0.5,
    hjust = 0.5
  ) +
  scale_fill_gradientn('Percentage of population vaccinated',
                       colours = rev(hcl.colors(10,"Green-Yellow")),
                       na.value = "grey90") +
  coord_map() +
  scale_x_continuous(breaks = NULL) + scale_y_continuous(breaks = NULL) +
  labs(x = "", y = "")  +
  ggtitle(paste0("Heatmap of vaccination rates across the country on end of Sep'21")) +
  theme(panel.background = element_blank(),
        plot.title = element_text(face = 'bold', size = 18))

})
```

### Trend shift infavor of Online channels influenced by the pandemic
```{r eruptions4, echo=FALSE}
renderPlot({
  ggplot(data=sales, aes(x=Order, y=Total, group=1)) +
  geom_line(arrow = arrow(), color = "blue") + ggtitle("Change in percentage of online sales over quarters") +  scale_x_discrete(name="Timeline", limits=vec) +
  geom_point() + labs(y = "%age of E-commerce retail sales") +
  theme(panel.background = element_blank()) + 
  geom_text(data = sales, aes(y = Total, label = paste(Total, "%", sep = "")), size=2.5, vjust=-1.2)
})
```

### We want to identify states that have shown better growth in Retail GDP from Q4'20 to Q1'21 and also states that have higher Retail demand in Q1'21 :

```{r eruptions2, echo=FALSE}

renderPlot({
  ggplot(gdp, aes(x = gdp2021Q1, y = growth_rate_Q1_Q4)) +
    geom_point(color = "darkblue", size = 1.5) +
    labs(x = "2021 Q1 GDP of Retail industry (Millions)",
         y = "Growth rate QoQ (2021Q1 vs. 2020Q4)",
         title = "Q1 GDP and growth rate of Retail industry") +
    theme_light() +
    geom_text(aes(label = `ï..state`), hjust = 0.5, vjust = 1.6)  +
    scale_y_continuous(labels = scales::percent,
                       breaks = seq(0.025, 0.055, by = 0.005)) +
    theme(plot.title = element_text(hjust = 0.5, size = 15)) +
    geom_point(
      data = gdphigh,
      aes(x = gdp2021Q1, y = growth_rate_Q1_Q4),
      colour = "red",
      size = 3
    )
})
```



